Practical Optional Types for Clojure

### 2015-02-24 formal-model.tex
classpe example

\smallsection{\classpe{}}
If \inpropenv {\propenv{}} {\isprop {\Union{\Keyword}{\Nil}} {\x{}}}
and
\inpropenv {\propenv{}} {\isprop {\Value{\Keyword}} {\path{\classpe{}}{\x{}}}}

>>>>

\smallsection{\classpe{}}
Consider the then branch of the following program:

\begin{minted}{clojure}
(fn [x :- (U nil Keyword)]
  (when (class x)
    ; here
    ))
\end{minted}

Intuitively, since \clj{(class x)} is a true value it must be a class, therefore
\clj{x} is an object (not \clj{nil}).

From checking the body with T-Abs we know {\isprop {\Union{\Keyword}{\Nil}} {\x{}}}
The test has the object {\path{\classpe{}}{\x{}}} by T-App.
By T-If we can assume
{\notprop {\falsy{}} {\path{\classpe{}}{\x{}}}} in the then branch.
Since the object of the first proposition {\x{}} is a prefix of the second proposition's object
{\path{\classpe{}}{\x{}}}, 
by T-Update 
$$
\inpropenv {\propenv{}} {\isprop {\update {\Union{\Keyword}{\Nil}} {\nottype {\falsy{}}} {\classpe{}}} {\x{}}}.
$$

This hits the sixth \classpe{} case of \updateliteral{}, which returns
$$
{\update {\Union{\Keyword}{\Nil}} {\Object} {\emptypath{}}}
$$
deriving \inpropenv {\propenv{}} {\isprop {\Keyword} {\x{}}} 
by \restrictliteral{} (figure~\ref{main:figure:restrictremove}).
